### 2024-10-03 esop25/cpj.tex
checkpoint

\section{Confidentiality Types}

@@

For example, given a message send such as:
$$
\xassign{\mx{x}{1}}{\secret{x} \ftimes \flip{x}}{2}
$$
we would assign the following type to denote that $\mx{x}{1}$ is dependent on
both $\sx{x}{2}$ and $\rx{x}{2}$:
$$
\mx{x}{1} : \setit{\sx{x}{2}, \rx{x}{2}}
$$
But in the case of 2-party reconstructive sharing:
$$
\xassign{\mx{x}{1}}{\secret{x} \fminus \flip{x}}{2}
$$
we want the type of $\mx{x}{1}$ to be an ``encrypted'' type
that is by itself independent of $\sx{x}{2}$. However, we also
want to track the dependence of $\mx{x}{2}$ and $\rx{x}{2}$.
In short, we assign $\mx{x}{1}$ a ``ciphertext type'':
$$
\mx{x}{1} : \cty{\rx{x}{2}}{\setit{\sx{x}{2}}}
$$
indicating syntactically that $\mx{x}{1}$ is an encryption of a
value dependent on $\sx{x}{2}$ with $\rx{x}{2}$ and as long as client
1 does not subsequently receive messages that are dependent on
$\rx{x}{2}$, the value $\sx{x}{2}$ remains hidden.

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and the
``evaluator'' obtains the wire value for its input via OT. The
evaluator shares the initial encrypted wire value of one of its input
secrets as a selection bit-- assuming that client 2 is the garbler and
client 1 is the evaluator, we can write this as:
$$
\xassign{\mx{x}{1}}{\mux{\secret{x}}{\neg\flip{x}}{\flip{x}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{x}{1}}{\OT{\secret{x}}{1}{\neg\flip{x}}{\flip{x}}}{2}}
   \toeq{\prog} \models \mx{x}{1} \eop \neg\sx{x}{2} \fplus \rx{x}{2}
\end{mathpar}
This allows us to assign the same ciphertext type to $\mx{x}{1}$ as
above-- that is, $\cty{\rx{x}{2}}{\setit{\sx{x}{2}}}$. Leveraging the
equivalence this way has been previously well-studied
\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
it is also important that we ensure that samples are used at most once
for encryption. As in previous work \cite{darais2019language} we can
use a form of type linearity to this end.

@@

Given a valid typing, we can leverage structural type information to
conservatively approximate the information accessible by the
adversary who has control over all corrupt clients in $C$ given
a partitioning $H,C$. In particular, if the adversary has access to an
encypted value and another value that is dependent on its one-time-pad
key-- i.e., access to values with types of the forms:
$$
\setit{\cty{\rx{x}{\cid}}{\ty}} \qquad \setit{\rx{x}{\cid}, \ldots} 
$$
then we can conservatively assume that the adversary can decrypt the
value of type $\ty$ and thus have access to those values.

>>>>

\section{Confidentiality Types}
\label{section-cpj}

@@

For example, given some string $y$ and a message send such as:
$$
\xassign{\mx{w}{1}}{\secret{w} \ftimes \flip{w}}{2}
$$
we would assign the following type to denote that $\mx{w}{1}$ is dependent on
both $\sx{w}{2}$ and $\rx{w}{2}$:
$$
\mx{w}{1} : \setit{\sx{w}{2}, \rx{w}{2}}
$$
But in the case of 2-party reconstructive sharing:
$$
\xassign{\mx{w}{1}}{\secret{w} \fminus \flip{w}}{2}
$$
we want the type of $\mx{w}{1}$ to be an ``encrypted'' type
that is by itself independent of $\sx{w}{2}$. However, we also
want to track the dependence of $\mx{w}{2}$ and $\rx{w}{2}$.
In short, we assign $\mx{w}{1}$ a ``ciphertext type'':
$$
\mx{w}{1} : \cty{\rx{w}{2}}{\setit{\sx{w}{2}}}
$$
indicating syntactically that $\mx{w}{1}$ is an encryption of a
value dependent on $\sx{w}{2}$ with $\rx{w}{2}$ and as long as client
1 does not subsequently receive messages that are dependent on
$\rx{w}{2}$, the value $\sx{w}{2}$ remains hidden.

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and the
``evaluator'' obtains the wire value for its input via OT. The
evaluator shares the initial encrypted wire value of one of its input
secrets as a selection bit-- assuming that client 2 is the garbler and
client 1 is the evaluator, we can write this as:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
This allows us to assign the same ciphertext type to $\mx{w}{1}$ as
above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
equivalence this way has been previously well-studied
\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
it is also important that we ensure that samples are used at most once
for encryption. As in previous work \cite{darais2019language} we can
use a form of type linearity to this end.

@@

Given a valid typing, we can leverage structural type information to
conservatively approximate the information accessible by the
adversary who has control over all corrupt clients in $C$ given
a partitioning $H,C$. In particular, if the adversary has access to an
encypted value and another value that is dependent on its one-time-pad
key-- i.e., access to values with types of the forms:
$$
\setit{\cty{\rx{w}{\cid}}{\ty}} \qquad \setit{\rx{w}{\cid}, \ldots} 
$$
then we can conservatively assume that the adversary can decrypt the
value of type $\ty$ and thus have access to those values.

### 2024-10-03 esop25/ipj.tex
checkpoint

\section{Integrity Types}

@@

The typing of protocols depends on a type judgement for expressions
of the form $\itj{\cid}{\be}{\ity{\cid}{V}}$, which records
messages and reveals $V$ used in the construction of $\be$.
In an assignment, this source is recorded in $\Delta$-- so for
example, letting:
$$
\prog\ \defeq\ (\xassign{\mx{x}{1}}{\secret{x} - \flip{x}}{2};\ \rvl{x} := \mx{x}{1})
$$
the following judgement is valid:
$$
\ipj{\toeq{\prog}}{\prog}{(\mx{x}{1} : \ity{2}{\varnothing}; \rvl{x} : \ity{1}{\mx{x}{1}})}
$$
More formally, we defined validity as follows:
\begin{definition}
  Given pre-processing predicate $\eqspre$ and protocol $\prog$, 
  we say $\ipj{\eqs}{\prog}{\Delta}$ is \emph{valid}
  iff it is derivable and $\eqspre \wedge \toeq{\prog} \models E$.
\end{definition}

In the above example, note that environment implicitly records
that $\rvl{x}$ depends on data from client 1, which in turn
depends on data from client 2. The rule for MAC
checking allows an overwrite of the dependency. We explore
the application of this in more detail in Section
\ref{section-examples}.

@@

In all cases, we want to determine the integrity level of computed
values based on assumed partitionins $H,C$. In the previous example,
we want to say that $\rvl{x}$ has low integrity if either client 1 or
2 is corrupt.  To this end we define \emph{security labelings}
$\seclev$ which are mappings from variables $x$ to security levels
$\latel$. By default, given a partitioning $H,C$ this labeling maps
messages to the integrity level of the owner.
\begin{definition}  
  Given $H,C$,
  define $\seclev_{H,C}$ such that $\seclev_{H,C}(\mx{w}{\cid} ) = \hilab$
  if $\cid \in H$  and $\lolab$
  otherwise for all $\mx{w}{\cid}$.
\end{definition}
During execution of a protocol, the integrity of messages and reveals
affected by values sent by other parties, as recorded in $\Delta$
obtained from its typing. We additionally need to determine
whether this impacts integrity, which we accomplish through
the inductively defined rewrite relation $\cheatj{\Delta}{H,C}{\seclev}$,
axiomatized in Figure \ref{fig-cheatj}.
Our main type safety result for integrity is then given as follows.
\begin{theorem}
  Given pre-processing predicate $\eqspre$ and protocol $\prog$ with
  $\views(\prog) = V$, if
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  and for all $H,C$ with $\cheatj{\Delta}{H,C}{\seclev}$ 
  we have $\seclev(x) = \hilab$ for all $x \in \houtputs$, then cheating
  is detectable in $\prog$.
\end{theorem}

>>>>

\section{Integrity Types}
\label{section-ipj}

@@

The typing of protocols depends on a type judgement for expressions
of the form $\itj{\cid}{\be}{\ity{\cid}{V}}$, which records
messages and reveals $V$ used in the construction of $\be$.
In an assignment, this source is recorded in $\Delta$-- so for
example, for some string $w$ letting:
$$
\prog\ \defeq\ (\xassign{\mx{w}{1}}{\secret{w} - \flip{w}}{2};\ \rvl{w} := \mx{w}{1})
$$
the following judgement is valid:
$$
\ipj{\toeq{\prog}}{\prog}{(\mx{w}{1} : \ity{2}{\varnothing}; \rvl{w} : \ity{1}{\mx{w}{1}})}
$$
More formally, we defined validity as follows:
\begin{definition}
  Given pre-processing predicate $\eqspre$ and protocol $\prog$, 
  we say $\ipj{\eqs}{\prog}{\Delta}$ is \emph{valid}
  iff it is derivable and $\eqspre \wedge \toeq{\prog} \models E$.
\end{definition}

In the above example, note that environment implicitly records
that $\rvl{w}$ depends on data from client 1, which in turn
depends on data from client 2. The rule for MAC
checking allows an overwrite of the dependency. We explore
the application of this in more detail in Section
\ref{section-examples}.

@@

In all cases, we want to determine the integrity level of computed
values based on assumed partitionins $H,C$. In the previous example,
we want to say that $\rvl{w}$ has low integrity if either client 1 or
2 is corrupt.  To this end we define \emph{security labelings}
$\seclev$ which are mappings from variables $x$ to security levels
$\latel$. By default, given a partitioning $H,C$ this labeling maps
messages to the integrity level of the owner.
\begin{definition}  
  Given $H,C$,
  define $\seclev_{H,C}$ such that $\seclev_{H,C}(\mx{w}{\cid} ) = \hilab$
  if $\cid \in H$  and $\lolab$
  otherwise for all $\mx{w}{\cid}$.
\end{definition}
During execution of a protocol, the integrity of messages and reveals
affected by values sent by other parties, as recorded in $\Delta$
obtained from its typing. We additionally need to determine
whether this impacts integrity, which we accomplish through
the inductively defined rewrite relation $\cheatj{\Delta}{H,C}{\seclev}$,
axiomatized in Figure \ref{fig-cheatj}.
Our main type safety result for integrity is then given as follows.
\begin{theorem}
  Given pre-processing predicate $\eqspre$ and protocol $\prog$ with
  $\views(\prog) = V$, if
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  and for all $H,C$ with $\cheatj{\Delta}{H,C}{\seclev}$ 
  we have $\seclev(x) = \hilab$ for all $x \in \houtputs$, then cheating
  is detectable in $\prog$.
\end{theorem}

### 2024-10-03 esop25/metalang.tex
checkpoint

We define a big-step evaluation relation $\redx$ in Figures
\ref{fig-metaprotexprsemantics} and \ref{fig-metaprotinstrsemantics}
for expressions and instructions, respectively.  In this definition we
write $e[\mv/y]$ and $\instr[\mv/y]$ to denote the substitution of
$\mv$ for free occurrences of $y$ in $e$ or $\instr$ respectively. The
rules are mostly standard.

@@

>>>>

We define a big-step evaluation relation $\redx$ in Figures
\ref{fig-metaprotexprsemantics} and \ref{fig-metaprotinstrsemantics}
for expressions and instructions, respectively.  In this definition we
write $e[\mv/y]$ and $\instr[\mv/y]$ to denote the substitution of
$\mv$ for free occurrences of $y$ in $e$ or $\instr$ respectively. The
rules are mostly standard. Note that we do not include an evaluation
rule for the form $\eqcast{e}{e}{\notg{\phi}}$ which only has
relevance for the type system and which we assume is erased from
programs prior to evaluation. We defer discussion of this form,
as well as the syntactic category $\notg{\phi}$, to the next
Section.

@@

Our first goal with a $\metaprot$ type theory is to define an
algorithmic system that is sound for both confidentiality and
integrity typings as defined in Sections \ref{section-cpj} and
\ref{section-ipj}. Returning to the example in Section
\ref{section-cpj}, note that the key equivalence of the $\ttt{mux}$
expression with a one-time-pad encryption we observe there
is not trivial. To ensure that this sort of equivalence
can be picked up by the type system, we introduce a
an annotation form that allows the programmer to provide
the needed hint.

Consider the following $\ttt{encode}$ function definition that
generalizes this encoding for any identifier $y$ and sender, receiver
pair $s,r$:
$$
\begin{array}{l}
\ttt{encode}(y,s,r) \{\\
\quad \xassign{\mx{y}{r}}{\mux{\secret{y}}{\neg\flip{y}}{\flip{y}}}{s};\\
\quad \eqcast{\mx{y}{r}}{\neg\sx{y}{s} \fplus \rx{y}{s}}\\
\}
\end{array}
$$

Furthermore, we want to minimize the amount of SMT solving needed for
type verification.  Our approach is to support pre- and post-condition
annotations on $\metaprot$ functions that can be independently
verified, allowing significant reduction of SMT overhead in
whole-program analysis.
